---
title: "network"
author: "Misha Ash"
date: "5/26/2018"
output: 
  flexdashboard::flex_dashboard:
    vertical_layout: fill
---

### a glimpse of caregiving relationships
    
I got caught in quite a data munging rabbit hole, but crawled out eventually.

Using Color Brewer, I attempted to build a scale of color vectors to modify the node colors but could not get it working. 

The arrows did not work for me either — I tried making them edges directed from the domain of caregivers to the codomain of clients.

    
### caregiving network

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(broom)
library(stringr)

library(flexdashboard)

library(networkD3)
library(igraph)
library(network)
library(sna)
library(visNetwork)
library(threejs)
library(ndtv)


options(digits = 3)
set.seed(144)

base_size <- 22
theme_set(theme_minimal(base_size = base_size))
```

```{r import, include = FALSE}
data <- read_csv("~/R/Data Viz/ntg-informavorous/057_net_data_anon 2.csv", col_names = TRUE,
                 col_types = list(col_character(), 
                                  col_character(), 
                                  col_character(),
                                  col_double(), 
                                  col_character(), 
                                  col_character()))[-1,] %>%
  as.data.frame() %>%
  select(CX, CG, TTL, "IHSS ?", "REL ?") %>%
  separate(CX, into = c("to_type", "to"), sep = "_") %>%
  separate(CG, into = c("from_type", "from"), sep = "_")

as.tibble(data)
```

```{r simple network, eval = FALSE, include = FALSE}
simpleNetwork(data[ , 1:2])
```

```{r network tests, eval = FALSE, include = FALSE}
net <- graph_from_data_frame(d = edges, vertices = nodes, directed = T) 
net

plot(net)

net <- simplify(net, remove.multiple = F, remove.loops = T) 

plot(net, edge.arrow.size=.9,vertex.label=NA)

l <- layout_in_circle(net)
plot(net, layout = l)

cut.off <- mean(nodes$TTL) 
net.sp <- delete_edges(net, E(net)[TTL<cut.off])
plot(net.sp) 

netm <- as_adjacency_matrix(net, attr="TTL", sparse=F)

palf <- colorRampPalette(c("gold", "dark orange")) 
heatmap(netm[1:50,50:1], Rowv = NA, Colv = NA, col = palf(100), 
        scale="none", margins=c(10,10) )

deg.dist <- degree_distribution(net, cumulative=T, mode="all")
plot( x=0:max(deg), y=1-deg.dist, pch=19, cex=1.2, col="orange", 
      xlab="Degree", ylab="Cumulative Frequency")

E(net)$width <- 1.5
plot(net, edge.color=c("dark red", "slategrey")[(E(net)$type=="hyperlink")+1],
      vertex.color="gray40", layout=layout_in_circle, edge.curved=.7)

visNetwork(nodes, edges, width="100%", height="400px", main="Network!")

```


```{r edges, include = FALSE}
net_edges <- data %>%
  mutate(domain = sort(.$from)) %>% 
  mutate(domain = as.numeric(factor(.$from)) -1) %>%
  
  # continue numbering of codomain taking domain cardinality into account (reproduceably)
  mutate(codomain = as.numeric(factor(.$to)) + (length(unique(.$from))) - 1) %>%
  # create alternatively labeled vector specifying relation type
  mutate(relation = if_else(`REL ?` == "N", "unrelated", "relatives")) %>%
  # remove unnecessary variables
  select(-`IHSS ?`, -`REL ?`)
  

```

```{r nodes, include = FALSE}

net_nodes <- data %>%
  mutate(codomain = str_c(to, to_type, sep="_")) %>%
  mutate(domain = str_c(from, from_type, sep="_")) %>%
  select(domain, codomain, TTL, `IHSS ?`) %>%
  mutate(waiver = if_else(`IHSS ?` == "N", "PC-HMK", "IHSS")) %>%
  select(-`IHSS ?`) %>%

#  unite(to_type, to, sep = "_")
  
  gather(type, id, -TTL, -waiver)  %>%
  separate(id, into = c("id", "role"), sep = "_") %>%
  mutate(property = str_c(role, waiver, sep = "_")) %>%
  select(-waiver, -type, -role) %>%
  
  group_by(property, id) %>%

  summarize(TTL = sum(TTL)) %>%
  arrange(id, property) %>%
  as.data.frame() %>%
  mutate(index = row_number() -1)


duplicates <- filter(net_nodes[duplicated(net_nodes$id), ]) %>%
  mutate(X = "MULT") %>%
  select(id, X)

net_nodes <- net_nodes %>%
  left_join(duplicates, by = "id") 

# replace NAs in X vector (to subsequently replace them with values from property)
net_nodes[is.na(net_nodes$X), 5] <- 0

net_nodes <- net_nodes %>%
  separate(property, into = c("role", "property"), sep = "_") %>%
  mutate(waiver = if_else(X == "MULT", "MULT", property)) %>%
  
  select(-index, -X, -property) %>%
  
  group_by(role, waiver, id) %>%

  summarize(TTL = sum(TTL)) %>%
  arrange(role, id) %>%
  as.data.frame() %>%
  mutate(index = row_number() -1) %>%
  mutate(group = str_c(role, waiver, sep = " "))

```

```{r edge weights, include = FALSE}

edge_width_scale <- function(x){(((x-min(x))/(max(x)-min(x))) + 1)*2}
x <- edge_width_scale(net_nodes$TTL)
```

```{r edge colors, include = FALSE}

## Specify colours for specific edges
# Find links to Valjean (11)
which(net_nodes == "domain", arr = TRUE)[1] - 1

edge_feature <-  which(net_edges == "relatives", arr = TRUE)[, 1]

# Create a colour vector
edge_colors <- ifelse(1:nrow(net_edges) %in% edge_feature, "#d73027", "#fee08b")

```

```{r node colors, include = FALSE}
# for colourScale parameter

node_colors <- c("#1b7837", "#5aae61", "#a6dba0", "#9970ab", "#762a83", "#fdb863")
```

```{r forceNetwork, echo = FALSE}
forceNetwork(Links = net_edges, Nodes = net_nodes, Source = "domain", Target = "codomain",
               NodeID = "index", Group = "group", linkWidth = x,
               linkColour = edge_colors, #colourScale = "node_colors",
               fontSize = 12, zoom = T, legend = T,
               Nodesize = 4, opacity = 0.9, charge = -8) #, 
              # width = 600, height = 400)
```


---




```{r export, eval = FALSE, include = FALSE}
nodes <- read.csv("Dataset1-Media-Example-NODES.csv", header=T, as.is=T)
links <- read.csv("Dataset1-Media-Example-EDGES.csv", header=T, as.is=T)
```


```{r eval = FALSE, include = FALSE}
library(scales)

net_nodes <- nodes %>%
  mutate(test = sort(.$id)) #%>% 

  summarize(TTL = sum(TTL))

net_nodes <- cbind(idn = factor(net_nodes$id, levels = net_nodes$id), net_nodes)

# net_nodes <- cbind(idnum = as.numeric(factor(net_nodes$id, levels = net_nodes$id)) -1, net_nodes)


```

```{r munging struggle residues, eval = FALSE, include = FALSE}

row.names(net_nodes[duplicated(net_nodes$id), ])


# identify which columns contain the values you might want to replace
data[ , 2:3 ]


# subset it with extended bracketing..
data[ , 2:3 ][ data[ , 2:3 ] == 4 ]
# ..those were the values you're going to replace

# now overwrite 'em with tens
data[ , 2:3 ][ data[ , 2:3 ] == 4 ] <- 10

# look at the final data
data






  mutate(node = if_else(X == "MULT", "MULT", property)) %>%
  select(id, TTL, index, node) %>%


  mutate(properties = (ifelse(X == "MULT", "MULT", .$property)))

net_nodes_TEST <- net_nodes %>%
  mutate(TEST = replace(.$property, 
                          list = row.names(net_nodes[duplicated(net_nodes$id), ]), 
                          values = "BOTH"))

P <- xor(row.names(net_nodes[duplicated(net_nodes$id), ]))

net_nodes[duplicated(net_nodes$id), 1] <- net_nodes[duplicated(net_nodes$id), "BOTH"]

Q <- filter(net_nodes[duplicated(net_nodes$id), ])

df[find.a, "X1"] <- df[find.a, "X2"]


  mutate(properties = ifelse(.$id == net_nodes[duplicated(net_nodes$id), 2],
                              "BOTH", property))


```

